---
title: "Annual Escapement and Goals"
author: "Madeline Jovanovich"
date: "5/28/2018"
output: html_document
---
For some unknown reason I started added harvest to this rmd, but that was a mistake. I need to make three htmls: 

1. escapement with goals 
2. harvest
3. hatchery

```{r libraries, message=FALSE, warning=FALSE, echo=FALSE}
library(leaflet)
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)
library(scales)
library(knitr)
library(ggthemes)
library(RColorBrewer)
library(colorspace)
library(scales)
library(viridis)
library(DT)
```

Need to change to KNB url once it is generated: **check on this** 
```{r data}
escapement_goals <- read.csv("~/R/Esc_goals_evaluated.csv", stringsAsFactors = FALSE) %>%
  arrange(SASAP.Region, LocationID, Species)
```

```{r Changing_Region_name_for_simplicity}
escapement_goals$SASAP.Region[escapement_goals$SASAP.Region == "Alaska Peninsula and Aleutian Islands"] <- "AK Pen. and Aleutians"

escapement_goals$types[escapement_goals$types == "inRiver"] <- "In-River"

escapement_goals$types[escapement_goals$types == "agreement"] <- "Agreement"
```


```{r Color Schemes, echo=FALSE, include=FALSE}
species_color <- c(chinook = "#33A02C", coho = "#A6CEE3", chum = "#1F78B4", sockeye = "#E31A1C", pink = "#FB9A99")
show_col(species_color)
```

####Arctic Region

Show there are no goals in the Arctic by adding a region to the data set: this will be displayed in the table and plot that shows goal by region
```{r Arctic Region, warning=FALSE}
Arctic <- data.frame("Arctic")
colnames(Arctic) <- "SASAP.Region"
```

####Year QA

```{r QA Data - Year Ranges and Complete Data}
YearQA <- escapement_goals %>%
  select(SASAP.Region, sampleYear) %>%
  group_by(SASAP.Region) %>%
  summarize(min(sampleYear),
            max(sampleYear))

#Bristol Bay is the only region that includes 2017, so let's remove it (causing issues anyways)

escapement_goals_QAed <- escapement_goals %>%
  filter(sampleYear < 2017) #This is done to ignore BB 2017 data, which is incomplete as of 5/2/18, but still included in cumulative escapement data

escapement_goals_QAedkable <- escapement_goals_QAed %>%
  select(SASAP.Region, sampleYear) %>%
  group_by(SASAP.Region) %>%
  summarize(min(sampleYear),
            max(sampleYear))
#Print a kable to show year ranges, and make note of this exclusion

eg_summary <- bind_rows(Arctic, escapement_goals_QAedkable) %>%
  arrange(SASAP.Region)
```

```{r kable showing Orginial data}
#Should make these tables include a list of goal types by region
kable(YearQA, caption = "Ranges of Years in Original Dataset")
```

**Missing most recent 2 years: notice that not all regions have the same max year. Should I only evaluate thorough 2014? Seems rather silly given that would exclude all of Bristol Bay's most recent large runs, as well as the past couple Kusko years of slight improvement.** 

```{r kable showing Year QA}
kable(escapement_goals_QAedkable, caption = "Ranges of Years after removing incomplete 2017 data")
```

#### Establishing proportions of achievement

This may need some work
```{r Establishing Proportions of Achievement}
#Need to create a kable of this
#Do I really want to remove NA values? If they are included in the df, I can show locations of annual counts with no goals
#that way, EG_results would retains NAs, percents would omit them

EG_results <- escapement_goals_QAed %>%
  filter(!is.na(MetLower) | !is.na(MetUpper)) %>%
  arrange(SASAP.Region, LocationID)

sumna <- function(x) {
    sum(x, na.rm = TRUE)
}

#Selecting a lot of columns here, but may eventually need to make a separate df: one with only a few pertinent columns for "percents", another with more data
percents <- EG_results %>%
    #select(SASAP.Region, LocationID, Species, sampleYear, MetLower, MetUpper) %>%
    select(SASAP.Region, LocationID, Species, sampleYear, annualCount, types, MetLower, MetUpper) %>%
    group_by(sampleYear, SASAP.Region, types, LocationID, Species) %>%
    summarise(countMetLower = sumna(MetLower), 
              countMetUpper = sumna(MetUpper), 
              notMetUpper = sumna(!MetUpper), 
              notMetLower = sumna(!MetLower),
              percentMetUpper = (sumna(MetUpper)/(sumna(MetUpper) + sumna(!MetUpper))),
              percentMetLower = (sumna(MetLower)/(sumna(MetLower) + sumna(!MetLower)))) %>%
    arrange(SASAP.Region, LocationID)

percents$metUnmetExceeded = ifelse(percents$countMetUpper > 0, "Exceeded", ifelse(percents$countMetLower > 0, "Met", "Unmet"))

plottingdf <- bind_rows(percents, Arctic) %>%
  arrange(SASAP.Region, LocationID)
```

#Summary Tables
```{r escapement_range, include = FALSE}
range <- escapement_goals %>%
  group_by(SASAP.Region, Species) %>%
  summarize (min_count = min(annualCount), max_count = max(annualCount), min_year = min(sampleYear), max_year = max(sampleYear)) %>%
   mutate_if(is.numeric, funs(round(., 0)))
datatable(range)
```

Need to learn how to count number of character observations... feel like I've done that before.
```{r location_count, include = FALSE}
locations <- escapement_goals %>% 
  group_by(SASAP.Region, Species) 
#%>%
 # summarize(count = count(unique(LocationID)) #this doesn't work, but this line shows what I would like to do 
datatable(locations)
```

Leaflet map included in M. Jovanovich's WDAFS talk 
```{r location_lat_lon, include=FALSE}
mean_esc <- escapement_goals %>%
  group_by(SASAP.Region, Lat, Lon, Species, LocationID) %>%
  summarize (meanby_sp = mean(annualCount)) %>%
   mutate_if(is.numeric, funs(round(., 0))) %>%
  arrange(SASAP.Region)
datatable(mean_esc)
#I want this to show the average species count by region, not by location
```

```{r leaflet, echo=FALSE}
markerIcon <- makeIcon(
  iconUrl = "~/R/icons/salmon.png",
  iconWidth = 25, iconHeight = 41,
  iconAnchorX = 12, iconAnchorY = 41
)
leaflet(escapement_goals) %>% 
  addTiles() %>% 
  addMarkers(~ Lon, ~ Lat, popup = ~ Species, icon = markerIcon, clusterOptions = markerClusterOptions())
```

```{r goal_by_regionI}
ggplot(data = plottingdf, aes(x = SASAP.Region, fill=factor(types), y=(..count..))) + 
    geom_bar() +
    labs(x = "",
         y = "Number of Goals", 
         title = "Escapement Goals by Region", 
         subtitle = "note: y-axis is an accumulation of all goals for all species over time") +
    ggtitle("Escapement Goals by Region") +
    theme_hc() +
    theme(aspect.ratio = 0.5,
          plot.background = element_rect(fill="gray96"), 
          legend.text = element_text(size=8), 
          legend.position = "right", 
          legend.background = element_rect(fill="gray96"), 
          panel.grid.major.y = element_line(color="gray93"),
          axis.ticks.x = element_line(color = "gray96")) +
    guides(fill=guide_legend("Goal Type"), legend.title=element_text(size=10)) +
    theme(axis.text.x = element_text(angle = 50, hjust = 0.85)) +
    scale_fill_brewer("Colors in Spectral", palette="Spectral", breaks = c("Agreement", "BEG", "In-River", "MT", "OEG", "SEG")) 
```

Note multiple_types created here
```{r goals_by_regionII, fig.align="center", echo=FALSE}
#Make this into pie charts
multiple_types <- subset(percents, SASAP.Region != "Copper River")
multiple_types <- subset(multiple_types, SASAP.Region != "Kotzebue")

ggplot(data = multiple_types, aes(x = sampleYear, fill=factor(types), y=(..count..))) + 
  geom_bar(width=1) +
    labs(x = "Year",
         y = "Number of Goals", 
         title = "Escapement Goals by Region") +
    #scale_y_continuous(labels = f) +
    scale_x_continuous(breaks = pretty_breaks(n = 10)) +
    theme_hc() +
    theme(plot.background = element_rect(fill="gray98"),
          legend.text = element_text(size=8), 
          legend.position = "none", 
          legend.background = element_rect(fill="gray98"),
          axis.text.x = element_text(angle = 50, hjust = 0.95),
          strip.background = element_rect(fill = "gray98"),
          panel.grid.major.y = element_line(color="gray93")) +
    guides(fill=guide_legend("Goal Type")) +
    scale_fill_brewer("Colors in Spectral", palette="Spectral") + 
    facet_wrap(~SASAP.Region, scales="free_y", ncol = 2)
```

```{r goals_by_species, fig.align="center", echo=FALSE}
ggplot(data = EG_results, aes(x = "", fill=factor(types))) + 
    geom_bar() +
    labs(x = "Year",
         y = "Number of Goals", 
         title = "Escapement Goals by Species") +
    #scale_x_continuous(breaks = pretty_breaks(n = 10)) +
    theme_hc() +
    theme(aspect.ratio = 0.2,
          plot.background = element_rect(fill="gray98"),
          legend.text = element_text(size=8), 
          legend.position = "right", 
          legend.background = element_rect(fill="gray98"),
          axis.text.x = element_text(angle = 50, hjust = 0.95),
          strip.background = element_rect(fill = "gray98"),
          panel.grid.major.y = element_line(color="gray93")) +
    guides(fill=guide_legend("Goal Type"), legend.title=element_text(size=10)) +
    scale_fill_brewer("Colors in Spectral", palette="Spectral") + 
    facet_wrap(~Species, ncol=1) + 
  coord_polar("y")
#This looks crazy right now, but really cool how it's plotting on a 3D plane! 
```
```{r goals_by_regionIII}
#Need to further look in how to make these; not working when I add facet_wrap 
ggplot(data = percents, aes(x = "", fill=factor(metUnmetExceeded))) + 
    geom_bar(width = 1) +
    labs(title = "Annual Goal Achievement, Statewide") +
    scale_y_continuous(labels=percent) + 
    guides(fill=guide_legend(title="")) + 
    theme_hc() +
  scale_fill_brewer(palette = "YlGnBu", direction = -1) +
  coord_polar("y") +
  theme(axis.text.x=element_blank()) + 
  facet_wrap(~SASAP.Region)
```

```{r statewide_achievement, fig.align="center", echo=FALSE}
ggplot(data = percents, aes(x = "", fill=factor(metUnmetExceeded), y=(..count..))) + 
    geom_bar(width = 1) +
    labs(title = "Annual Goal Achievement, Statewide") +
    scale_y_continuous(labels=percent) + 
    guides(fill=guide_legend(title="")) + 
    theme_hc() +
  scale_fill_brewer(palette = "YlGnBu", direction = -1) +
  coord_polar("y") +
  theme(axis.text.x=element_blank(), axis.text.y=element_blank())

```

#Escapement + Goal Bounds 

Be sure to include the Arctic here, to show lack of escapement and goals. 

Workflow: complete this task for one region, then adapt it to each region

1. Summary for each region
  * escapement by location? 
  * escapement by species 
2. Escapement over time for subset of locations
  * include a datatable of locations within each region
  * chose a couple obvious examples, therefore showing how to subset by location 


```{r}
Kvichak <- escapement_goals %>%
  filter(LocationID == "Kvichak River")

ggplot(data = Kvichak, aes(x = sampleYear, y = annualCount, color = Species)) + 
    geom_line(size = 1) +
      expand_limits(y=c(0, 1000000)) +
      scale_y_continuous(labels=comma) +
      scale_x_continuous(breaks = pretty_breaks(n = 10)) +
      labs(x = "Year", 
           y = "Annual Escapement", 
           title = "Kvichak River Sockeye") + 
      guides(fill=guide_legend(title="Species")) + 
      theme_hc() +
       theme(plot.background = element_rect(fill="gray98"),
          legend.text = element_text(size=8), 
          legend.position = "none", 
          legend.background = element_rect(fill="gray98"),
          strip.background = element_rect(fill = "gray98"),
          panel.grid.major.y = element_line(color="gray93"),
          axis.text.x = element_text(angle = 50, hjust = 0.85),
          plot.title = element_text(margin=margin(0,0,25,0))) +
      scale_color_manual(values = species_color) + 
geom_hline(yintercept = 2e+06)
```

##Alaska Peninsula and Aleutian Islands

```{r akpenaleut_locations}
akpenaleut_locations <- escapement_goals %>%
  group_by(LocationID) %>%
  select(LocationID) #why is this showing one line for each row of data instead of one line for each locationID? 
datatable(akpenaleut_locations)

```

```{r example_plot_with_theme}
ggplot(data = Kvichak, aes(x = sampleYear, y = annualCount, color = Species)) + 
    geom_line(size = 1) +
      expand_limits(y=c(0, 1000000)) +
      scale_y_continuous(labels=comma) +
      scale_x_continuous(breaks = pretty_breaks(n = 10)) +
      labs(x = "Year", 
           y = "Annual Escapement", 
           title = "Kvichak River Sockeye") + 
      guides(fill=guide_legend(title="Species")) + 
      theme_hc() +
       theme(plot.background = element_rect(fill="gray98"),
          legend.text = element_text(size=8), 
          legend.position = "none", 
          legend.background = element_rect(fill="gray98"),
          strip.background = element_rect(fill = "gray98"),
          panel.grid.major.y = element_line(color="gray93"),
          axis.text.x = element_text(angle = 50, hjust = 0.85),
          plot.title = element_text(margin=margin(0,0,25,0))) +
      scale_color_manual(values = species_color) + 
geom_hline(yintercept = 2e+06)
```


###Alaska Peninsula and Aleutian Islands by Species  

##Bristol Bay

###Bristol Bay by Species

##Chignik

###Chignik by Species 

##Cook Inlet

###Cook Inlet by Species

##Copper River 

###Copper River by Species 

##Kodiak

###Kodiak by Species 

##Kotzebue 

###Kotzebue by Species 

##Kuskokwim 

###Kuskokwim by Species 

```{r Kusko_all_species}
ggplot(joined, aes(x=Year, y=NumFish, fill=DataType)) +
    geom_bar(stat = "identity") + 
    labs (title = "Kuskokwim River Harvest and Escapement", 
          subtitle = "all species", 
          x = "Year", 
          y = "Number of Fish") + 
    scale_y_continuous(labels = comma, expand = c(0, 0)) +
    guides(fill=guide_legend(title="")) + 
    theme_hc() +
        theme(plot.background = element_rect(fill="gray98"),
          legend.text = element_text(size=8), 
          legend.position = "right", 
          legend.background = element_rect(fill="gray98"),
          strip.background = element_rect(fill = "gray98"),
          panel.grid.major.y = element_line(color="gray93"), 
          axis.text.x = element_text(angle = 50, hjust = 0.85))

Kuskojoined_summed <- joined %>% 
  group_by(Species, Year) %>% 
  summarize(sum_count = sum(NumFish))

ggplot(Kuskojoined_summed, aes(x=Year, y=sum_count)) + #color = #need to figure out how to get DataType to work
    geom_line(size = 1) + 
    labs (title = "Kuskokwim River Harvest and Escapement: Chinook", 
          x = "Year", 
          y = "Number of Fish") + 
    scale_y_continuous(labels = comma, expand = c(0, 0), limits = c(0, max(Chinook$NumFish)), breaks = pretty_breaks(n = 4)) +
    scale_x_continuous(breaks = pretty_breaks(n = 10)) +
    guides(color=guide_legend(title="")) + 
    theme_hc() +
        theme(plot.background = element_rect(fill="gray98"),
          legend.text = element_text(size=8), 
          legend.position = "right", 
          legend.background = element_rect(fill="gray98"),
          strip.background = element_rect(fill = "gray98"),
          panel.grid.major.y = element_line(color="gray93"),
          axis.text.x = element_text(angle = 50, hjust = 0.85),
          plot.title = element_text(margin=margin(0,0,25,0))) 
#Need to make this plot summed annual counts so there is one observation per year
```
```{r Kusko_Chinook}
Chinook <- subset(joined, Species == "chinook")

ggplot(Chinook, aes(x=Year, y=NumFish, fill=DataType)) +
    geom_bar(stat = "identity") + 
    labs (title = "Kuskokwim River Harvest and Escapement: Chinook", 
          x = "Year", 
          y = "Number of Fish") + 
    scale_y_continuous(labels = comma, expand = c(0, 0)) +
    scale_x_continuous(breaks = pretty_breaks(n = 10)) +
    guides(fill=guide_legend(title="")) + 
    theme_hc() +
        theme(plot.background = element_rect(fill="gray98"),
          legend.text = element_text(size=8), 
          legend.position = "right", 
          legend.background = element_rect(fill="gray98"),
          strip.background = element_rect(fill = "gray98"),
          panel.grid.major.y = element_line(color="gray93"), 
          axis.text.x = element_text(angle = 50, hjust = 0.85))

ggplot(Chinook, aes(x=Year, y=NumFish, color=DataType)) +
    geom_line(size = 1) + 
    labs (title = "Kuskokwim River Harvest and Escapement: Chinook", 
          x = "Year", 
          y = "Number of Fish") + 
    scale_y_continuous(labels = comma, expand = c(0, 0), limits = c(0, max(Chinook$NumFish)), breaks = pretty_breaks(n = 4)) +
    scale_x_continuous(breaks = pretty_breaks(n = 10)) +
    guides(color=guide_legend(title="")) + 
    theme_hc() +
        theme(plot.background = element_rect(fill="gray98"),
          legend.text = element_text(size=8), 
          legend.position = "right", 
          legend.background = element_rect(fill="gray98"),
          strip.background = element_rect(fill = "gray98"),
          panel.grid.major.y = element_line(color="gray93"),
          axis.text.x = element_text(angle = 50, hjust = 0.85),
          plot.title = element_text(margin=margin(0,0,25,0))) + 
  geom_line()
#I NEED TO MAKE A SUBSET OF KUSKOKWIM RIVER CHINOOK, CHOOSE ONE SYSTEM TO LOOK AT, PLOT THAT ESCAPEMENT GOAL ON TOP OF THESE PLOTS 
```
```{r Kusko_sockeye}
Sockeye <- subset(joined, Species == "sockeye")

ggplot(Sockeye, aes(x=Year, y=NumFish, fill=DataType)) +
    geom_bar(stat = "identity") + 
    labs (title = "Kuskokwim River Harvest and Escapement: Sockeye", 
          x = "Year", 
          y = "Number of Fish") + 
    scale_y_continuous(labels = comma, expand = c(0, 0), breaks = pretty_breaks(n = 6)) +
    scale_x_continuous(breaks = pretty_breaks(n = 10)) +
    guides(fill=guide_legend(title="")) + 
    theme_hc() +
        theme(plot.background = element_rect(fill="gray98"),
          legend.text = element_text(size=8), 
          legend.position = "right", 
          legend.background = element_rect(fill="gray98"),
          strip.background = element_rect(fill = "gray98"),
          panel.grid.major.y = element_line(color="gray93"), 
          axis.text.x = element_text(angle = 50, hjust = 0.85))

ggplot(Sockeye, aes(x=Year, y=NumFish, color=DataType)) +
    geom_line(size = 1) + 
    labs (title = "Kuskokwim River Harvest and Escapement: Sockeye", 
          x = "Year", 
          y = "Number of Fish") + 
    scale_y_continuous(labels = comma, expand = c(0, 0), limits = c(0, max(Sockeye$NumFish)), breaks = pretty_breaks(n = 6)) +
    scale_x_continuous(breaks = pretty_breaks(n = 10)) +
    guides(color=guide_legend(title="")) + 
    theme_hc() +
        theme(plot.background = element_rect(fill="gray98"),
          legend.text = element_text(size=8), 
          legend.position = "right", 
          legend.background = element_rect(fill="gray98"),
          strip.background = element_rect(fill = "gray98"),
          panel.grid.major.y = element_line(color="gray93"),
          axis.text.x = element_text(angle = 50, hjust = 0.85),
          plot.title = element_text(margin=margin(0,0,25,0)))
```
```{r Kusko_coho}
Coho <- subset(joined, Species == "coho")

ggplot(Coho, aes(x=Year, y=NumFish, fill=DataType)) +
    geom_bar(stat = "identity") + 
    labs (title = "Kuskokwim River Harvest and Escapement: Coho", 
          x = "Year", 
          y = "Number of Fish") + 
    scale_y_continuous(labels = comma, expand = c(0, 0), breaks = pretty_breaks(n = 5)) +
    scale_x_continuous(breaks = pretty_breaks(n = 10)) +
    guides(fill=guide_legend(title="")) + 
    theme_hc() +
        theme(plot.background = element_rect(fill="gray98"),
          legend.text = element_text(size=8), 
          legend.position = "right", 
          legend.background = element_rect(fill="gray98"),
          strip.background = element_rect(fill = "gray98"),
          panel.grid.major.y = element_line(color="gray93"), 
          axis.text.x = element_text(angle = 50, hjust = 0.85))

ggplot(Coho, aes(x=Year, y=NumFish, color=DataType)) +
    geom_line(size = 1) + 
    labs (title = "Kuskokwim River Harvest and Escapement: Coho", 
          x = "Year", 
          y = "Number of Fish") + 
    scale_y_continuous(labels = comma, expand = c(0, 0), limits = c(0, max(Coho$NumFish)), breaks = pretty_breaks(n = 5)) +
    scale_x_continuous(breaks = pretty_breaks(n = 10)) +
    guides(color=guide_legend(title="")) + 
    theme_hc() +
        theme(plot.background = element_rect(fill="gray98"),
          legend.text = element_text(size=8), 
          legend.position = "right", 
          legend.background = element_rect(fill="gray98"),
          strip.background = element_rect(fill = "gray98"),
          panel.grid.major.y = element_line(color="gray93"),
          axis.text.x = element_text(angle = 50, hjust = 0.85),
          plot.title = element_text(margin=margin(0,0,25,0)))
```
```{r Kusko_chum}
Chum <- subset(joined, Species == "chum")

ggplot(Chum, aes(x=Year, y=NumFish, fill=DataType)) +
    geom_bar(stat = "identity") + 
    labs (title = "Kuskokwim River Harvest and Escapement: Chum", 
          x = "Year", 
          y = "Number of Fish") + 
    scale_y_continuous(labels = comma, expand = c(0, 0), breaks = pretty_breaks(n = 5)) +
    scale_x_continuous(breaks = pretty_breaks(n = 10)) +
    guides(fill=guide_legend(title="")) + 
    theme_hc() +
        theme(plot.background = element_rect(fill="gray98"),
          legend.text = element_text(size=8), 
          legend.position = "right", 
          legend.background = element_rect(fill="gray98"),
          strip.background = element_rect(fill = "gray98"),
          panel.grid.major.y = element_line(color="gray93"), 
          axis.text.x = element_text(angle = 50, hjust = 0.85))

ggplot(Chum, aes(x=Year, y=NumFish, color=DataType)) +
    geom_line(size = 1) + 
    labs (title = "Kuskokwim River Harvest and Escapement: Chum", 
          x = "Year", 
          y = "Number of Fish") + 
    scale_y_continuous(labels = comma, expand = c(0, 0), limits = c(0, 3000000), breaks = pretty_breaks(n = 5)) +
    scale_x_continuous(breaks = pretty_breaks(n = 10)) +
    guides(color=guide_legend(title="")) + 
    theme_hc() +
        theme(plot.background = element_rect(fill="gray98"),
          legend.text = element_text(size=8), 
          legend.position = "right", 
          legend.background = element_rect(fill="gray98"),
          strip.background = element_rect(fill = "gray98"),
          panel.grid.major.y = element_line(color="gray93"),
          axis.text.x = element_text(angle = 50, hjust = 0.85),
          plot.title = element_text(margin=margin(0,0,25,0)))
```

##Norton Sound 

###Norton Sound by Species 

##Prince William Sound

###Prince William Sound by Species

##Southeast 

###Southeast by Species 

##Yukon 

###Yukon by Species 

