---
title: "Alaska Hatcheries: joining releases and returns"
author: "Madeline Jovanovich"
date: "3/13/2018"
output: 
  html_document: 
    highlight: tango
    theme: paper
    toc: yes
    toc_depth: 3
---


Pool/sum across years, split across regions for releases and returns, facet by species; additionally have a total all species
Release and returns faceted by species for each region 

Table: species, region, total release, total return, summed across years = total # of fish releases by species within a given region 

Table: total release and return summed across species and years by region 

Table: shows types of hatchery releases by release type


**TO DOs**

*   Were appropriate columns retained from each data set? 
        -   Should we have two joined data sets: one with all data in addition to the one created below? 
*   Retain released type, so data can be filtered 
*   Investigate NA and 0 values 
*   Investigate suspect values
*   Clean up plot axes

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = F, warning = F)
```

```{r, message=FALSE}
library(dplyr)
library(ggplot2)
library(scales)
library(tidyr)
library(knitr)
library(kableExtra)
```

###Reformatting data

Read in data from KNB, aggregate over region, year, and species, join together, and make a long dataframe for easier faceting.

```{r}
returns <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/knb.92369.1", stringsAsFactors = F)
#returns <- read.csv("~/R/Hatchery_Returns.csv", stringsAsFactors = F)

returns_species_region <- returns %>%
    group_by(SASAP.Region, Year, Species) %>%
    summarise(TotalReturn = sum(Catch))

releases <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Aaf27c17d-3aad-41e4-b1c0-57ac8bb8a022", stringsAsFactors = F)
colnames(releases)[colnames(releases) == "SASAPRegion"] <- "SASAP.Region"
colnames(releases)[colnames(releases) == "YearReleased"] <- "Year"

releases_species_region <- releases %>%
    group_by(SASAP.Region, Year, Species) %>%
    summarise(TotalReleased = sum(TotalReleased))


ret_rel <- left_join(releases_species_region, returns_species_region)

ret_rel_long <- ret_rel %>%
    gather(key = "ReleaseReturn", value = "NumFish", -SASAP.Region, -Year, -Species)
```
```{r results = 'asis', fig.align="center", echo=FALSE}
kable(ret_rel_long[c(1:3, 20:22, 40:42, 60:62),], caption = "Releases and Returns Data") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```


###Plotting by Species

```{r, echo=FALSE}
ret_rel_long_species <- ret_rel_long %>%
    group_by(Year, Species, ReleaseReturn) %>%
    summarize(TotalFish = sum(NumFish, na.rm = T))
```
```{r results = 'asis', echo=FALSE}
kable(ret_rel_long_species[c(1:3, 20:22, 40:42, 60:62),], caption = "Restructured Data for Plotting") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
```{r}
ggplot(ret_rel_long_species, aes(x = Year, y = TotalFish, color = Species, fill = Species)) + 
    geom_bar(stat = "identity") +
    scale_y_continuous(labels = comma) + 
    labs(title = "Statewide Summed Hatchery Releases and Returns", 
         y = "Total Fish", 
         x = "Year")
```
```{r}
ggplot(ret_rel_long_species, aes(x = Year, y = TotalFish, color = Species, fill = Species)) + 
    geom_bar(stat = "identity") +
    facet_wrap(~ReleaseReturn, scales = "free", nrow = 2) +
    scale_y_continuous(labels = comma) + 
    labs(title = "Hatchery Releases and Returns", 
         y = "Total Fish", 
         x = "Year")
```
```{r}
ggplot(ret_rel_long_species, aes(x = Year, y = TotalFish, color = ReleaseReturn, fill = ReleaseReturn)) + 
    geom_bar(stat = "identity") +
    scale_y_continuous(labels = comma) + 
    labs(title = "Hatchery Releases and Returns", 
         y = "Total Fish", 
         x = "Year") + 
    theme_minimal() + 
    facet_wrap(~Species, scales = "free_y", ncol=2)
```

```{r}
ggplot(ret_rel_long_species, aes(x = Year, y = TotalFish, color = ReleaseReturn)) + 
    geom_line(size=0.75) +
    scale_y_continuous(labels = comma) + 
    labs(title = "Hatchery Releases and Returns", 
         y = "Total Fish", 
         x = "Year") + 
    theme_minimal() + 
    facet_wrap(~Species, scales = "free_y", ncol=2)
```

###Plotting by Region

```{r}
ret_rel_long_region <- ret_rel_long %>%
    group_by(Year, SASAP.Region, ReleaseReturn) %>%
    summarize(TotalFish = sum(NumFish, na.rm = TRUE))
```
```{r results = 'asis', echo=FALSE}
kable(ret_rel_long_region[c(1:3, 20:22, 40:42),], caption = "Restructured Data for Plotting") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
```{r}
ggplot(ret_rel_long_region, aes(x = Year, y = TotalFish, color = SASAP.Region, fill = SASAP.Region)) + 
    geom_bar(stat = "identity") +
    scale_y_continuous(labels = comma) + 
    labs(title = "Hatchery Releases and Returns", 
         y = "Total Fish", 
         x = "Year")
```
```{r}
ggplot(ret_rel_long_region, aes(x = Year, y = TotalFish, color = SASAP.Region, fill = SASAP.Region)) + 
    geom_bar(stat = "identity") +
    facet_wrap(~ReleaseReturn, scales = "free", nrow = 2) +
    scale_y_continuous(labels = comma) + 
    labs(title = "Hatchery Releases and Returns", 
         y = "Total Fish", 
         x = "Year")
```
```{r, fig.height=8, fig.width=8}
ggplot(ret_rel_long_region, aes(x = Year, y = TotalFish, color = ReleaseReturn, fill = ReleaseReturn)) + 
    geom_bar(stat = "identity") +
    scale_y_continuous(labels = comma) + 
    labs(title = "Hatchery Releases and Returns", 
         y = "Total Fish", 
         x = "Year") + 
    theme_minimal() + 
    facet_wrap(~SASAP.Region, scales = "free_y", ncol=2)
```
```{r, eval=FALSE}
ggplot(ret_rel, aes(color = ReleaseReturn, fill = ReleaseReturn)) + 
    geom_line(x = Year, y = TotalReleased) +
    geom_line(x = Year, y = TotalReturn) +
    scale_y_continuous(labels = comma) + 
    labs(title = "Hatchery Releases and Returns", 
         y = "Total Fish", 
         x = "Year") + 
    theme_minimal() + 
    facet_wrap(~SASAP.Region, scales = "free_y", ncol=2)
```

To do on the following plot: 

-   NA values still appearing
```{r, fig.height=6, fig.width=6}
#NEED TO REVISIT PLOTTING WITH TWO AXES
ggplot(ret_rel_long_region, aes(x = Year, y = TotalFish, color = ReleaseReturn)) + 
    geom_line(size=0.75) +
    scale_y_continuous(labels = comma) + 
    labs(title = "Hatchery Releases and Returns", 
         y = "Total Fish", 
         x = "Year") + 
    theme_minimal() + 
    facet_wrap(~SASAP.Region, scales = "free_y", ncol=2)

```
```{r, eval=FALSE, include=FALSE}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  scale_y_continuous(
    "mpg (US)", 
    sec.axis = sec_axis(~ . * 1.20, name = "mpg (UK)")
  )
```