---
title: "Hatchery Release Inquiry"
author: "Madeline Jovanovich"
date: "3/12/2018"
output: 
  html_document: 
    highlight: tango
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Hatchery releases:

I have found that the releases I previously plotted does not include a PWS region, but rather a Central region = PWS and other regions are not broken out of their broad-scale groupings that are designated by ADF&G.

```{r}
NPAFC_releases <- read.csv(file='~/Hatchery/Hatchery_ReleasesAK.csv')
unique(NPAFC_releases$ReportingArea)
```

So Prince William Sound can only be looked at as part of the Central Region. Let's take a look at this other file that was found in [Drive](https://drive.google.com/open?id=1kGnGS2cDp-bVHhe92TrjRNPKKbxYcTpU).

```{r}
ADFG_releases <- read.csv("~/Hatchery/Hatchery_Releases_geo.csv")
```

```{r, message=FALSE}
library(dplyr)
colnames(ADFG_releases) <- c('ReleaseYear', 'ReleaseTotal', 'Agency', 'SiteDistrict', 'SiteRegion', 'Species', 
                              'FacilityQuadrant', 'Facility', 'Flag', 'StateProv', 'Country')
AKrel <- ADFG_releases %>%
    filter(StateProv == "Alaska")

library(knitr)
kable(AKrel[1:8,])
```

We can see that there are many details in this dataset

*   96 Reporting Agencies
*   1810 Facilities
*   13 Facility Quandrants (that may line up to regions, or portions of regions)

```{r}
unique(AKrel$Agency)
unique(AKrel$Facility)
unique(AKrel$FacilityQuadrant)
```

In summary, we have two files:

1. NPAFC-reported hatchery releases, only have generalized Central region **without PWS broken out**
2. ADF&G-reported hatchery releases, with facility and quandrant info, **may require manually pairing facilities to regions**

So we can look in to using this file, I just need information on it's origin (I believe it is from Rich, May 2017) and any associated notes. 

```{r, message=FALSE, warning=FALSE}
library(ggplot2)
Central_releases <- NPAFC_releases %>%
    filter(ReportingArea == "Central") %>%
    arrange(Year)
ggplot(Central_releases, aes(x=Year, y=Release_1000s, color=Species)) +
    geom_point() + 
    ggtitle("Central Region Hatchery Releases", subtitle = "NPAFC data") + 
    facet_wrap(~Species)
ggplot(AKrel, aes(x=ReleaseYear, y=ReleaseTotal, color=Species)) + 
    geom_point() + 
    ggtitle("Hatchery Releases from Other File", subtitle = "from other file, assume to be obtained from ADF&G") +
    facet_wrap(~Species)
```